---
title: "Santa Cruz Monarch Movement Dashboard"
page-layout: full
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(data.table)
library(knitr)

dashboard_summary <- readRDS(
  "data/dashboard_summary.rds"
)

tag_summary <- readRDS(
  "data/tag_summary.rds"
)

tag_meta <- readRDS(
  "data/tag_meta.rds"
)
```

This dashboard provides an interactive summary of Blu+ telemetry detections, movement episodes, and overwintering site use for western monarchs radio tagged in Santa Cruz, California, USA.

The dashboard is designed to help visualize detection histories, movement paths, site use, visit and foray episodes, and broad behavioral classifications through time and space.

## How to Use This Dashboard

Use the navigation bar at the top of the page to move between the dashboard views:

- all detections,
- detections by individual tag,
- all movement episodes,
- episodes by individual tag.

### Detection Explorer

The detection explorer shows detection-level movement records. It includes:

- site boundary polygons,
- detection points,
- detection-step lines,
- behavior classifications,
- episode-type point symbols,
- dashed transition lines between new episodes or dates.

### Detections by Tag

The tag-specific detection pages show detection records for one monarch tag at a time. These pages are useful for reviewing the movement history of individual tagged monarchs.

### Episode Explorer

The episode explorer summarizes movement episodes, including visits, forays, episode start and end points, and transitions among sites.

### Episodes by Tag

The tag-specific episode pages show episode histories for individual monarchs. These pages are useful for reviewing site use, visit behavior, forays, and movement sequences by individual tag.

## Study Summary

```{r}
#| echo: false

summary_table <- data.table(
  Metric = c(
    "Telemetry detections",
    "Movement episodes",
    "Tracked monarchs"
  ),
  Value = c(
    format(
      dashboard_summary$total_points,
      big.mark = ","
    ),
    format(
      dashboard_summary$n_episodes,
      big.mark = ","
    ),
    format(
      dashboard_summary$n_tags,
      big.mark = ","
    )
  )
)

kable(
  summary_table,
  format = "html",
  align = c(
    "l",
    "r"
  )
)
```

## Study Period

```{r}
#| echo: false
#| results: asis

cat(
  paste0(
    "**",
    format(
      dashboard_summary$first_detection,
      "%d-%b-%Y"
    ),
    " to ",
    format(
      dashboard_summary$last_detection,
      "%d-%b-%Y"
    ),
    "**"
  )
)
```

## Data Included

The dashboard currently includes:

- `r format(dashboard_summary$total_points, big.mark = ",")` telemetry detections
- `r format(dashboard_summary$n_episodes, big.mark = ",")` movement episodes
- `r format(dashboard_summary$n_tags, big.mark = ",")` individually tracked monarchs
- overwintering site boundary polygons
- detection-level movement records
- visit and foray episode summaries
- behavior classifications:
  - migration
  - premigration
  - winter
  
## Data Included

The dashboard currently includes:

- `r format(dashboard_summary$total_points, big.mark = ",")` telemetry detections
- `r format(dashboard_summary$n_episodes, big.mark = ",")` movement episodes
- `r format(dashboard_summary$n_tags, big.mark = ",")` individually tracked monarchs
- overwintering site boundary polygons
- detection-level movement records
- visit and foray episode summaries
- behavior classifications:
  - migration
  - premigration
  - winter

## Example Tag Summaries

The table below shows the ten tags with the greatest number of detections.

```{r}
#| echo: false

tag_meta_dt <-
  as.data.table(
    tag_meta
  )

tag_summary_dt <-
  as.data.table(
    tag_summary
  )

example_tags <-
  tag_summary_dt[
    ,
    .(
      tag_id,
      tracking_period,
      n_detections,
      n_episodes,
      n_visits,
      n_forays
    )
  ][
    tag_meta_dt[
      ,
      .(
        tag_id,
        Tagging_Date,
        Individual_ID,
        Sex,
        Wing_Cond,
        Mass_g,
        Tagging_Site,
        OE
      )
    ],
    on = "tag_id"
  ][
    order(-n_detections)
  ][
    1:min(10, .N)
  ]

setnames(
  example_tags,
  old = c(
    "tag_id",
    "tracking_period",
    "n_detections",
    "n_episodes",
    "n_visits",
    "n_forays",
    "Tagging_Site",
    "Sex",
    "Mass_g",
    "Wing_Cond",
    "OE"
  ),
  new = c(
    "Tag ID",
    "Tracking Period",
    "Detections",
    "Episodes",
    "Visits",
    "Forays",
    "Tagging Site",
    "Sex",
    "Mass (g)",
    "Wing Condition",
    "OE"
  ),
  skip_absent = TRUE
)

kable(
  example_tags[
    ,
    .(
      `Tag ID`,
      `Tracking Period`,
      Sex,
      `Tagging Site`,
      `Detections`,
      `Episodes`,
      `Visits`,
      `Forays`,
      `Mass (g)`,
      `Wing Condition`,
      OE
    )
  ],
  format = "html",
  digits = 1
)
```
