---
title: "Santa Cruz Monarch Movement Dashboard"
---

```{r}
#| echo: false
#| warning: false
#| message: false

library(data.table)

dashboard_summary <- readRDS(
  "data/dashboard_summary.rds"
)

tag_summary <- readRDS(
  "data/tag_summary.rds"
)
```

Interactive exploration of monarch movement patterns, detection histories, and movement episodes from the Santa Cruz Monarch Telemetry Study.

## Study Summary

```{r}
#| echo: false

dashboard_summary
```

## Available Explorers

### Detection Explorer

Explore individual monarch detections through time and space.

Features:

- Interactive map
- Site boundaries
- Behavioral states
- Individual tag histories
- Detection-level records

### Episode Explorer

Explore movement episodes among sites.

Features:

- Movement trajectories
- Episode start and end locations
- Behavioral classifications
- Visit and foray episodes
- Site transitions

## Example Tag Summaries

```{r}
#| echo: false

tag_summary[
  ,
  .(
    tag_id,
    tracking_period,
    n_detections,
    n_episodes,
    n_visits,
    n_forays,
    primary_site
  )
][
  order(-n_detections)
][
  1:min(10, .N)
]
```

## Data Included

The dashboard currently includes:

- `r format(dashboard_summary$total_points, big.mark = ",")` telemetry detections
- `r format(dashboard_summary$n_episodes, big.mark = ",")` movement episodes
- `r format(dashboard_summary$n_tags, big.mark = ",")` individually tracked monarchs
- Site boundary polygons
- Behavior classifications:
  - Migration
  - Premigration
  - Winter

## Study Period

```{r}
#| echo: false
#| results: asis

cat(
  paste0(
    "**",
    format(
      dashboard_summary$first_detection,
      "%d-%b-%Y"
    ),
    " to ",
    format(
      dashboard_summary$last_detection,
      "%d-%b-%Y"
    ),
    "**"
  )
)
```
